---
title: 'EDS 241: Assignment 4'
author: "Paloma Cartwright"
date: "`r Sys.Date()`"
output: 
    pdf_document:
    toc: false
    number_sections: no
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
librarian::shelf("AER", "stargazer", "here", "tidyr", "tidyverse", "dplyr","stringr", "janitor", "huxtable",
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "readxl", "estimatr", "patchwork",
           "car")


```

\noindent This question will ask you to estimate the price elasticity of demand for fresh sardines across 56 ports located in 4 European countries with monthly data from 2013 to 2019. The data are contained in the file EU_sardines.csv, which is available on Gauchospace.  

Each row in the data file is a combination of port location (where the fish is landed and sold) in a given year and month. You can ignore the fact that the sample is not balanced (the number of monthly observations varies across ports).  

For the assignment, you will need the following variables: year, month, country, port (port where sardines are landed and sold), price_euro_kg (price per kg in €), and volume_sold_kg (quantity of sardines sold in kg). In the questions below, I use log() to denote the natural logarithm. 

```{r}

data <- read_csv(here("EU_sardines.csv"))

```

(a)  Estimate a bivariate regression of log(volume_sold_kg) on log(price_euro_kg). What is the price elasticity of demand for sardines? Test the null hypothesis that the price elasticity is equal to -1.

```{r}
log_data <- data %>% 
  mutate(log_vol = log(volume_sold_kg), 
         log_price = log(price_euro_kg))

model1 <- lm_robust(log_price ~ log_vol, data = log_data)
summary(model1)
price_elas <- model1$coefficients[[2]]

# linearHypothesis(model1, c("log_vol=-1"), white.adjust = "hc2")
```

The price elasticity of demand for sardines is `r price_elas`.

(b) Like in Lecture 8 (see the IV.R script), we will use wind_m_s as an instrument for log(price_euro_kg). To begin, estimate the first-stage regression relating log(price_euro_kg) to wind_m_s. Interpret the estimated coefficient on wind speed.  Does it have the expected sign? Also test for the relevance of the instrument and whether it is a “weak” instrument by reporting the proper F-statistic. 

```{r}
fs1 <- lm(formula = log_price ~ wind_m_s, data = log_data)
summary(fs1)
```

The coefficient on wind speed is `r fs1$coefficients[[2]]`. On avergae, when wind speeds increase by 1 meter per second, the log price of sardines per kilogram increases by `r fs1$coefficients[[2]]` euros. This does have the expected sign because when wind speed increases, it is presumed to be harder to catch fish which means there is less supply so price goes up. 

```{r}
lH1 <- linearHypothesis(fs1, c("wind_m_s=0"), white.adjust = "hc2")

f_stat <- lH1$F[2]
```

The f-statistic for this linear hypothesis is `r f_stat` and this is not a weak instrument because it is greater than the threshold.

(c) Estimate the TSLS estimator of the price elasticity of demand for sardines using wind_m_s as an instrument for log(price_euro_kg). What is the estimated price elasticity of demand for sardines?  

```{r}
tsls1 <- ivreg(log_vol ~ log_price | wind_m_s, data = log_data)
summary(tsls1)

price_elas2 <- tsls1$coefficients[[2]]
```

The estimated price elasticity for sardines is `r price_elas2`.

(d) Repeat the exercise in (c), but include fixed effects for each year, month, and country.  [Hint: you can use the command “as.factor(country) + as.factor(year) + as.factor(month)” to the ivreg function in R]. Report the estimated price elasticity of demand and the F-statistic testing for relevant and non-weak instruments.  

```{r}

tsls2 <- ivreg(log_vol ~ log_price + as.factor(country) + as.factor(year) + as.factor(month) | wind_m_s, data = log_data)
summary(tsls2)
price_elas3 <- tsls2$coefficients[[2]]


```

